---
title: "1.8 — Price Elasticity — Practice Problems (Solutions)"
author: "ECON 306 - Fall 2020"
date: "Wednesday, September 9, 2020"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      fig.retina=3,
                      fig.align="center")
library(tidyverse)
library(ggrepel)
```

The demand for monthly cell phone plans is given by:

$$q_D =  50-0.5p$$

### 1. Write the *inverse* demand function.

Simply take the demand function and solve it for $p$:

$$\begin{align*}
	q_D&=50-0.5p\\
	q_D+0.5p&=50\\
	0.5p&=50-q_D\\
	p&=100-2q_D\\
\end{align*}$$

The vertical intercept (“choke price”) is \$100 and the slope is $-2$.

### 2. Calculate the price elasticity of demand when the price is $10. Is this relatively elastic or inelastic?

First we need to find $q_D$ at \$10. 

$$\begin{align*}
	q_D&=50-0.5(10)\\
	q_D&=50-5\\
	q_D&=45\\
\end{align*}$$

Now we have the three ingredients to calculate elasticity at \$10:
	 
$$\begin{align*}
	\epsilon_D &=\frac{1}{slope} \times \frac{p}{q_D}	\\
	\epsilon_D &=\frac{1}{-2} \times \frac{10}{45}	\\
	\epsilon_D &=-0.5 \times 0.22\\
	\epsilon_D &=-0.11\\
\end{align*}$$

The demand is relatively inelastic, as $|\epsilon_D|<1$ 

### 3. Calculate the price elasticity of demand when the price is $70. Is this relatively elastic or inelastic?

First we need to find $q_D$ at \$70. 

$$\begin{align*}
	q_D&=50-0.5(70)\\
	q_D&=50-35\\
	q_D&=15\\
	\end{align*}$$

We already have the slope (since the demand is a straight line), so now we can simply plug into the elasticity formula:

$$\begin{align*}
	\epsilon_D &=\frac{1}{slope} \times \frac{p}{q_D}	\\
	\epsilon_D &=\frac{1}{-2} \times \frac{70}{15}\\
	\epsilon_D &=-0.5 \times 4.67\\
	\epsilon_D &\approx -2.33\\
\end{align*}$$

The demand is relatively elastic, as $|\epsilon_D|>1$

### 4. At what price is demand unit elastic $(\epsilon=-1)$?

$$\begin{align*}
		\epsilon_D &=\frac{1}{slope} \times \frac{p}{q_D} & & \text{Formula for elasticity}\\
			-1&=-0.5 \times \frac{p}{q_D} & & \text{Set } \epsilon_D \text{ equal to }-1\\
			-1&=-0.5 \times \frac{p}{(50-0.5p)} & & \text{Plug in demand function for } q_D\\
			-1(50-0.5p)&=-0.5p & & \text{Multiply by term in parentheses}\\
		0.5p-50&=-0.5p & & \text{Distribute the }-1 \\
		-50&=-p & & \text{Add } 0.5p\\
		p&=\$50 & & \text{Divide by }-50\\
\end{align*}$$

### 5. Calculate the total revenue at $10.

The total revenue is:

$$\begin{align*}
R&=pq\\
R&=(\$10)(45)\\
R&=\$450\\
\end{align*}$$

### 6. Calculate the total revenue at $70.

The total revenue is:

$$\begin{align*}
R&=pq\\
R&=(\$70)(15)\\
R&=\$1,050\\
\end{align*}$$

### 7. Calculate the total revenue at the price you found for question 4.

That price was $p=\$50$. At this price, we need to find the quantity demanded. We can use the demand function: 

$$\begin{align*}
	q_D&=50-0.5p\\
	q_D&=50-0.5(50)\\
	q_D&=50-25\\
	q_D&=25\\	
	\end{align*}$$

Now that we have price and quantity, revenue is:

$$\begin{align*}
	R& = pq\\
	R&=(\$50)(25)\\	
	R&=\$1,250\\
	\end{align*}$$

This is where revenue is maximized.

```{r}
demand = function(x){100-2*x}

ggplot(data.frame(x=c(0,10)), aes(x=x))+
  stat_function(fun=demand, geom="line", size=2, color = "blue")+
  geom_label(aes(x=35,y=demand(35)), color = "blue", label="Demand", size = 5)+
  
  # line at 15,70
  geom_segment(aes(x=0,xend=15, y=70, yend=70), linetype="dashed")+
  geom_segment(aes(x=15,xend=15, y=70, yend=0), linetype="dashed")+
  
  # line at 25,50
  geom_segment(aes(x=0,xend=25, y=50, yend=50), linetype="dashed")+
  geom_segment(aes(x=25,xend=25, y=50, yend=0), linetype="dashed")+
  
  # line at 45,10
  geom_segment(aes(x=0,xend=45, y=10, yend=10), linetype="dashed")+
  geom_segment(aes(x=45,xend=45, y=10, yend=0), linetype="dashed")+
  
  geom_label_repel(data = tibble(x=15, y=70),
                   aes(x=x, y=y),
                  color="purple",
                  label=expression(paste(epsilon == -2.33, ": Elastic; R = $1,050")),
                  seed = 4 )+
  geom_label_repel(data = tibble(x=25, y=50),
                   aes(x=x, y=y),
                  color="purple",
                  label=expression(paste(epsilon == -1, ": Unit Elastic; R = $1,250")),
                  seed = 4 )+
  geom_label_repel(data = tibble(x=45, y=10),
                   aes(x=x, y=y),
                  color="purple",
                  label=expression(paste(epsilon == -0.11, ": Inelastic; R = $450")),
                  seed = 4 )+
  scale_x_continuous(breaks=seq(0,50,5),
                     limits=c(0,50),
                     expand=expand_scale(mult=c(0,0.1)))+
  scale_y_continuous(breaks=seq(0,100,10),
                     limits=c(0,100),
                     expand=expand_scale(mult=c(0,0.1)),
                     labels = function(x){paste("$", x, sep="")})+
  labs(x = "Quantity of x",
       y = "Price of x")+
  theme_classic(base_family = "Fira Sans Condensed", base_size=20)
```