---
title: "2.6 — Long Run Industry Equilibrium — Practice Problems"
author: "Solutions"
date: "ECON 306 - Fall 2020"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=F,
                      warning=F,
                      message=F)
library(tidyverse)
library(ggrepel)
library(mosaic)
```


The supply of bicycle rentals in a small town is given by:

$$q_S=10p-200$$

## 1. Find the inverse supply function.

---

Solve the supply function for $p$:

$$\begin{aligned}
	q_S&=10p-200 & & \text{The supply function}\\
	q_S-10p&=-200 & & \text{Subtracting }10p\\
	-10p&=-200-q_S & & \text{Subtracting }q_S\\
	p &=20+\frac{1}{10}q_S & & \text{Dividing by }-10\\	
	\end{aligned}$$

---

## 2. What is the price elasticity of supply at a price of \$25.00?

---

First we use the supply function and find $q_S$ when $p=25$: 

$$\begin{aligned}
			q_S&=10p-200 & & \text{The supply function}\\
			q_S&=10(25)-200 & & \text{Plugging in 25}\\
			q_S&=250-200 & & \text{Multiplying}\\
			q_S&=50 & & \text{Subtracting}\\
			\end{aligned}$$

Now we know the price, quantity, and slope, so we can use the elasticity formula.

$$\begin{aligned}
	\epsilon_S &=\frac{1}{slope} \times \frac{p}{q_S} & & \text{The formula for elasticity}\\
	\epsilon_S &=\cfrac{1}{\frac{1}{10}} \times \frac{25}{50} & & \text{Plugging in}\\
	\epsilon_S &=10 \times 0.5 & & \text{Simplifying}\\
	\epsilon_S &=5 & & \text{Multiplying}\\
	\end{aligned}$$

This is relatively elastic.

---

## 3. What is the price elasticity of supply at a price of \$50.00?

---

First we use the supply function and find $q_S$ when $p=50$: 

$$\begin{aligned}
			q_S&=10p-200 & & \text{The supply function}\\
			q_S&=10(50)-200 & & \text{Plugging in 25}\\
			q_S&=500-200 & & \text{Multiplying}\\
			q_S&=300 & & \text{Subtracting}\\
			\end{aligned}$$

Now we know the price, quantity, and slope, so we can use the elasticity formula. 

$$\begin{aligned}
	\epsilon_S &=\frac{1}{slope} \times \frac{p}{q_S} & & \text{The formula for elasticity}\\
	\epsilon_S &=\cfrac{1}{\frac{1}{10}} \times \frac{50}{300} & & \text{Plugging in}\\
	\epsilon_S &=10 \times \frac{50}{300} & & \text{Simplifying}\\
	\epsilon_S &=1.667 & & \text{Multiplying}\\
	\end{aligned}$$

This is also relatively elastic, but less elastic than at the lower price.

```{r, fig.retina=3}

supply<-function(x){20+0.1*x}
points<-tribble(
  ~x, ~y, ~label,
  50, 25, "E=5",
  300, 50, "E=1.67"
)

ggplot(data.frame(x=c(0,10)), aes(x=x))+
  stat_function(fun=supply, geom="line", size=2, color = "red")+
  geom_label(aes(x=700,y=supply(700)), color = "red", label="Supply", size = 6)+
    scale_x_continuous(breaks=seq(0,1000,100),
                     limits=c(0,1000),
                     expand=expand_scale(mult=c(0,0.1)))+
  scale_y_continuous(breaks=seq(0,100,10),
                     limits=c(0,100),
                     expand=expand_scale(mult=c(0,0.1)),
                     labels = function(x){paste("$", x, sep="")})+
  geom_segment(x=0,y=25,xend=50,yend=25,linetype="dotted",size=1)+
  geom_segment(x=50,y=0,xend=50,yend=25,linetype="dotted",size=1)+
  
  geom_segment(x=0,y=50,xend=300,yend=50,linetype="dotted",size=1)+
  geom_segment(x=300,y=0,xend=300,yend=50,linetype="dotted",size=1)+
  
  geom_point(data = points, aes(x=x,y=y), color="purple", size =3)+
  geom_label_repel(data = points, aes(x=x, y=y, label=label), color="purple", size =4)+
  labs(x = "Quantity of x",
       y = "Price of x")+
  ggthemes::theme_pander(base_family = "Fira Sans Condensed", base_size=14)
```