---
title: "4.4 — Factor Markets — Practice Problems"
author: "Solutions"
date: "ECON 306 - Fall 2020"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=F,
                      warning=F,
                      message=F)
library(tidyverse)
library(ggrepel)
library(mosaic)
```

Carl’s Coal Mining operates in a remote area. Because of its location, it has monopsony power in the local labor market for miners. Its marginal revenue product of labor is
$$MRP_L = 400-5L$$
where $L$ is the total number of miners. The labor supply curve of local miners is
$$w = 5L-50$$
where $w$ is the wage (in $1000's per miner).

### a. Write a function for the marginal cost of labor.

---

If we know the (inverse) labor supply function (which we are given), we can simply double the slope to find the marginal cost of labor:

$$MC(L)=10L-50$$

---

### b. What quantity of workers will the mine hire, and what wage will it pay its workers?

---

The optimal quantity of labor to hire for a firm is where its marginal revenue product is equal to the marginal cost of labor:

$$\begin{aligned}
MRP_L&=MC(L)\\
400-5L&=10L-50\\
400&=15L-50\\
450&=15L\\
30&=L^*\\
\end{aligned}$$

The firm has monopsony power, so it faces the entire market supply of labor. For $L^*$ number of workers, it can pay the lowest wages workers are willing to accept for that quantity, i.e. the labor supply function.

$$\begin{aligned}
w&=5L-50\\
w&=5(30)-50\\
w^*&=100\\
\end{aligned}$$

---

### c. What would the quantity of workers be, and what would the wage be, if there was competition among other local mines for labor?

---

If this was a competitive labor market, with no monopsony power, the firm would be a price-taker of labor, i.e. the supply of labor it faces would be perfectly elastic at the market-determined wage. It would set its marginal revenue product equal to the market wage and hire the quantity of workers for which those values are equal.

$$\begin{aligned}
MRP_L &= w\\
400-5L &= 5L-50\\
400&=10L-50\\
450&=10L\\
45&=L_c\\
\end{aligned}$$

Plug this quantity into the (inverse) labor supply function to the find the market wage:

$$\begin{aligned}
w&=5L-50\\
w&=5(45)-50\\
w&=225-50\\
w_C&=175\\
\end{aligned}$$

---

### d. Sketch a graph of this market, and be sure to label all of your findings (and show the Deadweight Loss) from Parts A-C.

---

```{r}
mrp_l=function(q){400-5*q}
supply_l=function(q){5*q-50}
MC_l=function(q){10*q-50}
dwl_l<-tribble(
  ~x, ~y,
  30, 250,
  30, 100,
  45, 175
)
ggplot(data.frame(x=c(0,10)), aes(x=x))+
  geom_polygon(data=dwl_l, aes(x=x,y=y), fill="black", alpha=0.5)+
    stat_function(fun=MC_l, geom="line", size=2, color = "darkred")+
  geom_label(aes(x=35,y=MC_l(35)), color = "darkred", label="MC(L)", size = 4)+
    stat_function(fun=supply_l, geom="line", size=2, color = "red")+
  geom_label(aes(x=35,y=supply_l(35)), color = "red", label="Supply(L)", size = 4)+
  stat_function(fun=mrp_l, geom="line", size=2, color = "blue")+
  geom_label(aes(x=15,y=mrp_l(15)), color = "blue", label=expression(MRP[L]), size = 4)+
  
  geom_segment(aes(x = 0, xend = 30, y = 100, yend=100), linetype="dotted", size=1)+
  geom_segment(aes(x = 30, xend = 30, y = 250, yend=0), linetype="dotted", size=1)+

  geom_segment(aes(x = 0, xend = 45, y = 175, yend=175), linetype="dotted", size=1)+
  geom_segment(aes(x = 45, xend = 45, y = 175, yend=0), linetype="dotted", size=1)+

    geom_label(aes(x=5,y=100), label=expression(w[m]), size = 4)+
    geom_label(aes(x=5,y=175), label=expression(w[C]), size = 4)+

      geom_label(aes(x=30,y=50), label=expression(L[m]), size = 4)+
    geom_label(aes(x=45,y=50), label=expression(L[C]), size = 4)+

    scale_x_continuous(breaks=seq(0,50,5),
                     limits=c(0,50),
                     expand=expand_scale(mult=c(0,0.1)))+
  scale_y_continuous(breaks=seq(0,400,50),
                     limits=c(0,400),
                     labels=scales::dollar,
                     expand=expand_scale(mult=c(0,0.1)))+
  #scale_colour_manual("Curves", values = line_colors, guide=F)+
  labs(y = "Wage (w)",
       x = "Quantity of Labor (L)")+
  ggthemes::theme_pander(base_size=12)
```